---
title: "Filtering to significant loci from GLM"
author: "Authors: Emma Strand; emma.strand@gmgi.org"
output:
  github_document: default
  pdf_document:
    keep_tex: yes
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r}
library(ggplot2)
library(tidyverse)
```

## Load data 

```{r}
load("/projects/gmgi/Fisheries/epiage/haddock/RData_Round2/df_all2.RData")
df_all2 <- df
rm(df)

sig_col_names <- c("scaffold", "start", "term", "df1", "df2", "F.ratio", "Chisq", "p.value")
stats_col_names <- c("scaffold", "start", "Max_Rhat", "n_eff")

# Function to read and rename columns of a single file
read_and_rename_sig <- function(file) {
  df <- read.table(file, header = TRUE)  # Read the text file
  colnames(df) <- sig_col_names  # Rename columns
  return(df)
}

read_and_rename_stats <- function(file) {
  df <- read.table(file, header = TRUE)  # Read the text file
  colnames(df) <- stats_col_names  # Rename columns
  return(df)
}
```

## df100 filtered 4 data

```{r}
# Define the path where your text files are located (switch this to run df100 or df)
#file_path <- "/projects/gmgi/Fisheries/epiage/haddock/GLM/df100_filtered4/seq2/"
file_path <- "/projects/gmgi/Fisheries/epiage/haddock/GLM/df_filtered4/"

# List all text files in the directory
sig_list_AL <- list.files(path = file_path, pattern = "sig_age_length_*", full.names = TRUE)
sig_list_age <- list.files(path = file_path, pattern = "sig_age_*", full.names = TRUE)
stats_list_AL <- list.files(path = file_path, pattern = "stats_age_length_*", full.names = TRUE)
stats_list_age <- list.files(path = file_path, pattern = "stats_age_*", full.names = TRUE)

# Read and rename columns for the first 12 files
datalist_sig_AL <- lapply(sig_list_AL, read_and_rename_sig)
datalist_sig_age <- lapply(sig_list_age, read_and_rename_sig)
datalist_stats_AL <- lapply(stats_list_AL, read_and_rename_stats)
datalist_stats_age <- lapply(stats_list_age, read_and_rename_stats)

# Combine all data frames into one and view first 6 rows with head()
df_sig_AL <- bind_rows(datalist_sig_AL); head(df_sig_AL)
df_sig_age <- bind_rows(datalist_sig_age); head(df_sig_age)
df_stats_AL <- bind_rows(datalist_stats_AL); head(df_stats_AL)
df_stats_age <- bind_rows(datalist_stats_age); head(df_stats_age)
```

## Filter sig df 

```{r}
# Function to filter columns and export the result
filter_sig <- function(df) {
  
  # Get the name of the input dataframe
  df_name <- deparse(substitute(df))
  
  # Create the Loc column, filter by p.value, and store in df_filtered
  df <- df %>% unite(Loc, c("scaffold", "start"), sep = " ", remove = FALSE)
  df_filtered <- df %>% filter(p.value < 0.050)
  
  # Print the number of unique Loc values in the original and filtered data frames
  cat("Unique locations in original data:", length(unique(df$Loc)), "\n")
  cat("Unique locations in filtered data:", length(unique(df_filtered$Loc)), "\n")
  
  # Create the new name for the filtered dataframe
  new_name <- paste0(df_name, "_filtered")
  
  # Assign the filtered dataframe to the new name in the global environment
  assign(new_name, df_filtered, envir = .GlobalEnv)
  
  # Inform the user about the new dataframe
  cat("Filtered dataframe exported as:", new_name, "\n")
  
}

## apply to dfs
filter_sig(df_sig_AL)
filter_sig(df_sig_age)
```

## Filtering stats df 

```{r}
# Function to filter columns and export the result
filter_stats <- function(df) {
  
  # Get the name of the input dataframe
  df_name <- deparse(substitute(df))
  
  # Create the Loc column, filter by p.value, and store in df_filtered
  df <- df %>% unite(Loc, c("scaffold", "start"), sep = " ", remove = FALSE)
  df_filtered <- df %>% filter(Max_Rhat < 1.010 & n_eff > 1000)
  
  # Print the number of unique Loc values in the original and filtered data frames
  cat("Unique locations in original data:", length(unique(df$Loc)), "\n")
  cat("Unique locations in filtered data:", length(unique(df_filtered$Loc)), "\n")
  
  # Create the new name for the filtered dataframe
  new_name <- paste0(df_name, "_filtered")
  
  # Assign the filtered dataframe to the new name in the global environment
  assign(new_name, df_filtered, envir = .GlobalEnv)
  
  # Inform the user about the new dataframe
  cat("Filtered dataframe exported as:", new_name, "\n")
  
}

## apply to dfs
filter_stats(df_stats_AL)
filter_stats(df_stats_age)
```

7-5-2026 notes for Max_Rhat < 1.010 & n_eff > 1000: 
Unique locations in original data: 635281 
Unique locations in filtered data: 308581 
Filtered dataframe exported as: df_stats_AL_filtered 
Unique locations in original data: 642346 
Unique locations in filtered data: 406711 
Filtered dataframe exported as: df_stats_age_filtered

3-24-2026 notes 
(Max_Rhat > 1.010 & n_eff < 2000) - OG incorrect filtering: 
Unique locations in original data: 635281 
Unique locations in filtered data: 213933 
Filtered dataframe exported as: df_stats_AL_filtered 
Unique locations in original data: 642346 
Unique locations in filtered data: 416414 
Filtered dataframe exported as: df_stats_age_filtered 

Max_Rhat < 1.010 & n_eff > 2000 - correct filtering but conservative:
Unique locations in original data: 635281 
Unique locations in filtered data: 182416 
Filtered dataframe exported as: df_stats_AL_filtered 
Unique locations in original data: 642346 
Unique locations in filtered data: 218667 
Filtered dataframe exported as: df_stats_age_filtered 


Joining those two dfs for 100% 

```{r}
# ### df100 filtered4 age only (20,323 sites; 7,761 sites)
# df100_f4_age_list <- inner_join(df_stats_age_filtered, df_sig_age_filtered, 
#                             by = c("Loc", "scaffold", "start")) %>% dplyr::select(Loc) %>% distinct()
# 
# df_sig_age_filtered %>% filter(Loc %in% df100_f4_age_list$Loc) %>% dplyr::select(Loc, p.value) %>% 
#   write.csv("/projects/gmgi/Fisheries/epiage/haddock/GLM/df100_filtered4/seq2/df100_f4_age_sig-7-05-2026.csv")
# 
# ### df100 filtered4 age + length (7,731 sites; 4,630 sites)
# df100_f4_agelength_list <- inner_join(df_stats_AL_filtered, df_sig_AL_filtered,
#                             by = c("Loc", "scaffold", "start")) %>% dplyr::select(Loc) %>% distinct()
# 
# df_sig_AL_filtered %>% filter(Loc %in% df100_f4_agelength_list$Loc) %>% dplyr::select(Loc, p.value) %>% 
#   write.csv("/projects/gmgi/Fisheries/epiage/haddock/GLM/df100_filtered4/seq2/df100_f4_AL_sig-7-05-2026.csv")
# 
# df100_f4_age_final <- df_all2 %>% right_join(., df100_f4_age_list, by = c("Loc"))
# df100_f4_agelength_final <- df_all2 %>% right_join(., df100_f4_agelength_list, by = c("Loc"))
# 
# save(df100_f4_age_final, file = "/projects/gmgi/Fisheries/epiage/haddock/GLM/df100_filtered4/seq2/df100_f4_age_final-7-05-2026.RData")
# save(df100_f4_agelength_final, file = "/projects/gmgi/Fisheries/epiage/haddock/GLM/df100_filtered4/seq2/df100_f4_agelength_final-7-05-2026.RData")
```

Joining those two dfs for 90% 

Error to troubleshoot: df_sig_age_filterd has multiple rows.. some loci went twice? Probably GLM array script issue.

```{r}
length(unique(df_sig_age_filtered$Loc)) ## 192,151 total; 124,836 sites post both filters
length(unique(df_sig_AL_filtered$Loc)) ## 140,055 total; 47,373 sites post both filters

### df filtered4 age only (124,836 sites; 66,191 sites)
df_f4_age_list <- inner_join(df_stats_age_filtered, df_sig_age_filtered, 
                            by = c("Loc", "scaffold", "start")) %>% dplyr::select(Loc) %>% distinct()

df_sig_age_filtered %>% filter(Loc %in% df_f4_age_list$Loc) %>% dplyr::select(Loc, p.value) %>% 
  write.csv("/projects/gmgi/Fisheries/epiage/haddock/GLM/df_filtered4/df_f4_age_sig-7-05-2026.csv")

### df filtered4 age + length (47,373 sites; 3-24-2026 40,251 sites)
df_f4_agelength_list <- inner_join(df_stats_AL_filtered, df_sig_AL_filtered,
                            by = c("Loc", "scaffold", "start")) %>% dplyr::select(Loc) %>% distinct()

df_sig_AL_filtered %>% filter(Loc %in% df_f4_agelength_list$Loc) %>% dplyr::select(Loc, p.value) %>% 
  write.csv("/projects/gmgi/Fisheries/epiage/haddock/GLM/df_filtered4/df_f4_AL_sig-7-05-2026.csv")

df_f4_age_final <- df_all2 %>% right_join(., df_f4_age_list, by = c("Loc"))
df_f4_agelength_final <- df_all2 %>% right_join(., df_f4_agelength_list, by = c("Loc"))

save(df_f4_age_final, file = "/projects/gmgi/Fisheries/epiage/haddock/GLM/df_filtered4/df_f4_age_final-7-05-2026.RData")
save(df_f4_agelength_final, file = "/projects/gmgi/Fisheries/epiage/haddock/GLM/df_filtered4/df_f4_agelength_final-7-05-2026.RData")
```












